---
title: "Sparse Grids"
author: Ahyan Panjwani
date: August 11th, 2020
output: html_notebook
---

I want to evaluate the function
$$
f(x, y) = \frac{1}{|0.5 - x^4 - y^4| + 0.1}
$$
which is not particularly _well-behaved_ in that it has a discontinuity due to the absolute value, and
it does not disappear at the boundaries. Evaluating the function through classical sparse grids is insufficient
because ot requires too many points to capture the ridge. Adaptive sparse grids are better suited for this
problem. This write-up demonstrates the potential of adaptive sparse grids.

## Brute Force

First, I directly evaluate the function over the unit square, $[0, 1]^2$ using $100^2$ points.

```{r warning=FALSE, message=FALSE, fig.align='center'}

library('plotly')

data.classic <- read.csv('/home/ahyan/Dropbox/Housing Market and Leverage Cycle/Code/src/Sparse/CSG.csv', header = FALSE)

fig1 <- plot_ly(data.classic, x = ~V1, y = ~V2, z = ~V3, marker = list(color = ~V3, colorscale = 'Portland', showscale = TRUE, size = 3))

fig1 <- fig1 %>% layout(title = 'Direct Evaluation')

fig1 <- fig1 %>% layout(scene = list(xaxis = list(title = 'x'),
                                     yaxis = list(title = 'y'),
                                     zaxis = list(title = 'f(x, y)')))

fig1

```

## Classical Sparse Grids

Now I evaluate the function on sparse grids of level 14. In two dimensions, that means almost 280,000 grid points.

```{r warning=FALSE, message=FALSE, fig.align='center'}
fig2 <- plot_ly(data.classic, x = ~V1, y = ~V2, z = ~V4, marker = list(color = ~V4, colorscale = 'Portland', showscale = TRUE, size = 3))

fig2 <- fig2 %>% layout(title = 'Classical Sparse Grid')

fig2 <- fig2 %>% layout(scene = list(xaxis = list(title = 'x'),
                                     yaxis = list(title = 'y'),
                                     zaxis = list(title = 'f(x, y)')))

fig2
```

The error between brute force and sparse grids, calculated as the $L_2$ norm is 0.0889.

$$
L_2 = \frac{1}{N} \bigg(\sum_{i = 1}^N |f(x_i) - u(x_i)|^2\bigg)^\frac{1}{2}
$$

## Adaptive Sparse Grids

Finally, I evaluate the function with adaptive sparse grids with level 5. In two dimensions, the algo starts
with 257 bounds (boundary included) and iterates until the error is less than 0.1 which happens with 6571
grid points and the error is 0.0897.

```{r warning=FALSE, message=FALSE, fig.align='center'}
data.adaptive <- read.csv('/home/ahyan/Dropbox/Housing Market and Leverage Cycle/Code/src/Sparse/ASG.csv', header = FALSE)

fig3 <- plot_ly(data.adaptive, x = ~V1, y = ~V2, z = ~V4, marker = list(color = ~V4, colorscale = 'Portland', showscale = TRUE, size = 3))

fig3 <- fig3 %>% layout(title = 'Adaptive Sparse Grids')

fig3 <- fig3 %>% layout(scene = list(xaxis = list(title = 'x'),
                                     yaxis = list(title = 'y'),
                                     zaxis = list(title = 'f(x, y)')))

fig3
```


## Life Cycle Problem

Basic
```{r warning=FALSE, message=FALSE, fig.align='center'}

data.life <- read.csv('/home/ahyan/Dropbox/Housing Market and Leverage Cycle/Code/src/Sparse/Life.csv', header = FALSE)

fig4 <- plot_ly(data.life, x = ~V1, y = ~V2, z = ~V3, marker = list(color = ~V3, colorscale = 'Portland', showscale = TRUE, size = 3))

fig4 <- fig4 %>% layout(title = 'Direct')

fig4 <- fig4 %>% layout(scene = list(xaxis = list(title = 'e'),
                                     yaxis = list(title = 'x'),
                                     zaxis = list(title = 'V(e, x)')))

fig4

```



```{r warning=FALSE, message=FALSE, fig.align='center'}

fig5 <- plot_ly(data.life, x = ~V1, y = ~V2, z = ~V4, marker = list(color = ~V4, colorscale = 'Portland', showscale = TRUE, size = 3))

fig5 <- fig5 %>% layout(title = 'Adaptive')

fig5 <- fig5 %>% layout(scene = list(xaxis = list(title = 'e'),
                                     yaxis = list(title = 'x'),
                                     zaxis = list(title = 'V(e, x)')))

fig5
```

With Nelder-Mead
```{r warning=FALSE, message=FALSE, fig.align='center'}

data.lifeB <- read.csv('/home/ahyan/Dropbox/Housing Market and Leverage Cycle/Code/src/Sparse/LifeB.csv', header = FALSE)

fig6 <- plot_ly(data.lifeB, x = ~V1, y = ~V2, z = ~V3, marker = list(color = ~V3, colorscale = 'Portland', showscale = TRUE, size = 3))

fig6 <- fig6 %>% layout(title = 'Direct')

fig6 <- fig6 %>% layout(scene = list(xaxis = list(title = 'e'),
                                     yaxis = list(title = 'x'),
                                     zaxis = list(title = 'V(e, x)')))

fig6

```
