---
title: "Life Cycle Consumption and Leverage"
author: "Ahyan Panjwani"
date: "6/30/2020"
output: 
  ioslides_presentation:
    mathjax: "http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    fontsize: 11pt
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Updates I

- Life cycle model of consumption and housing
- Households have the option to stay current, refinance, buy/sell, default
- Live for $T$ periods, leave a bequest
- Buy some housing at birth 
- Select mortgage from the credit surface offered
    - Endogenous leverage

## Updates II

- State space becoming very large:
    1. assets $a \in A$
    2. productivity $e \in E$ 
    3. housing quantity $q \in Q$
    4. $(m, OLTV) \in \mathcal{C}$
    5. $LTV \in L$ (contempranoeous loan-to-value)
    
$$
\text{state space = } A \times E \times Q \times \mathcal{C} \times L \subset \mathbb{R}^5
$$

- No FICO and rental market for now

## End of Life: Bequest

$$
\small
\begin{align*}
V(T, a, q, (m, OLTV), LTV, e; p) & = \max_{a'} u(c, q) \\
& + \beta \mathbb{E}V(T + 1, b'; p')\\
\end{align*}
$$

$$
\small
\begin{align*}
&c + a' + x(m, OLTV, q; p) = (1 + r)a + we\\
&b' = (1 + r) a' + (1 - LTV')\times p' \times q'\\
&p' = \Gamma(p)
\end{align*}
$$


where 

1. $x(m, OLT, q; p)$: the mortgage payment
2. $b'$: the expected bequest
3. $p' = \Gamma(p)$: rule for forming expectation of home price

## Choices Abound: Stay Current

$$
\small
\begin{align*}
&V^C(t, a , q, (m, OLTV), LTV, e; p) = \max_{a'} u(c, q) \\
& + \beta\mathbb{E}V(t + 1, a', q, (m, OLTV), LTV', e'; p')\\
&\\
& c + a' + x(m, OLTV, q; p) = (1 + r) a + we\\
& p' = \Gamma(p)\\
& e' \sim \mathbb{P}(e'|e)
\end{align*}
$$

where

1. $e' \sim \mathbb{P}(e'|e)$: Markov chain; AR(1) discretized a la Tauchen (1986)


## Choices abound: Refinance
$$
\small
\begin{align*}
& V^R(t, a, q, (m, OLTV), LTV, e; p) = \max_{a', (\overline{m}, \overline{OLTV}) \in \mathcal{C}} u(c, q) \\
& + \beta \mathbb{E} V(t + 1, a', q', (\overline{m}, \overline{OLTV}), \overline{LTV}', e'; p')\\
\end{align*}
$$
$$
\small
\begin{align*}
c + a' + x(\overline{m}, \overline{OLTV}, q; p) & = (1 + r)a + we \\
                                                & + (\overline{OLTV} - LTV)pq
\end{align*}
$$
where

1. $(\overline{m}, \overline{OLTV}) \in \mathcal{C}$: the new mortgage the agent selects from the surface
2. $\overline{LTV}$: contemporaneous LTV for the new mortgage
3. $(\overline{OLTV} - LTV)pq$: cash-out refinance windfall


## Choices abound: Buy/Sell

$$
\small
\begin{align*}
& V^S(t, a, q, (m, OLTV), LTV, e; p) = \max_{a', q', (\overline{m}, \overline{OLTV}) \in \mathcal{C}} u(c, q') \\
& + \beta \mathbb{E}V(t + 1, a', q', (\overline{m}, \overline{OLTV}), \overline{LTV}', e'; p')
\end{align*}
$$

$$
\small
\begin{align*}
c + a' + x(\overline{m}, \overline{OLTV}, q'; p) + (1 - \overline{OLTV})pq' & \\
= (1 + r)a + we + (1 - LTV)pq\\
\end{align*}
$$
where
1. $(1 - \overline{OLTV})pq'$: new downpayment
2. $(1 - LTV)pq$: equity windfall from sales

## Chocies abound: Default

$$
\small
\begin{align*}
& V^D(t, a, q, (m, OLTV), LTV, e; p) = \max_{a', q', (\overline{m}, \overline{OLTV}) \in \mathcal{C}} u(c, q) \\
& + \beta \mathbb{E}V(t + 1, a', q', (\overline{m}, \overline{OLTV}), \overline{LTV}', e'; p')
\end{align*}
$$
$$
\small
\begin{align*}
c + a' + x(\overline{m}, \overline{OLTV}, q'; p) + (1 - \overline{OLTV})pq' & \\
= (1 + r)a + we\\
\end{align*}
$$
where

1. **No** equity extraction
2. Live in same housing for on period

## Discrete Choice
$$
\small
\begin{equation*}
    V(t, a, q, (m, OLTV), LTV, e; p) = \max
    \begin{cases}
    V^C & stay \: current\\
    V^R & refinance \\
    V^S & buy/sell\\
    V^D & default
\end{cases}
\end{equation*}
$$

## Buying at Birth

$$
\small
\begin{align*}
V(0, a, e; p) &= \max_{a', q, (m, OLTV) \in \mathcal{C}} u(c, q) \\
              & + \beta\mathbb{E}V(1, a', q, (m, OLTV), LTV', e'; p')
\end{align*}
$$
$$
\small
\begin{align*}
c + a' + x(m, OLTV, q; p) + (1 - OLTV)pq = a + we 
\end{align*}
$$
where

1. $(1 - OLTV)pq$: downpayment

## Grids

- $A$: 20 points
- $Q$: 20 points
- $\mathcal{C}$: 26 points
- $L$: 10 points (i.e. mortgage term of 10 years)
- $E$: 9 points
- $T$: 10 points (i.e. 10 cohorts)

For each cohort: 

$$
\small
\begin{align*}
& A \times Q \times \mathcal{C} \times L \times E\\
\Rightarrow & 20 \times 20 \times 26 \times 10 \times 9\\
\Rightarrow & \sim 1 \: \text{million nodes / cohort}
\end{align*}
$$

## Model Logic
 - Given prices, lender develops credit surface
 - Agents optimize given their state variables, credit surface, and prices
 - House price, $p$, adjusts s.t. housing mkt clears
 
$$
\begin{align*}
Q_D(p) &= \int_{h \in H} q^h(p) = \overline{Q}_S
\end{align*}
$$
where $H$ is the state space, $\overline{Q}_S$ is fixed supply of housing

 - Each $h \in H$ represents a different household
 

## Computation

 - State space is very large
 - Use parallel within cohorts
 - Use nested loops for more efficiency
 - Precomputation of integrals
 - Starting to use *sparse grids* to reduce nodes
    + This is preliminary at this stage
    
<br>
All runs local

 - $\texttt{Intel Core i7-9750, 2.60GHz x 6 x 2}$
 - $\texttt{Ubuntu (Linux) 20.04; Java 11}$
 - Avg core temperature: 87 Celsius (critical at 100$^\circ$C)




## Computation: Results
+-----------------+--------------------------------------------+---------------+
|Model            |Computation                                 |Run-time       |
+=================+============================================+===============+
|C                |Parallel                                    | < 1m          |
+-----------------+--------------------------------------------+---------------+
|C, R             |Parallel (nested)                           | ~ 1m          |
+-----------------+--------------------------------------------+---------------+
|C, R, S          |Parallel (nested)                           | ~ 21m         |
+-----------------+--------------------------------------------+---------------+
|C, R, S, D       |Serial                                      | ~ 3.5h        |
+-----------------+--------------------------------------------+---------------+
|C, R, S, D       |Parallel                                    | ~ 57m         |
+-----------------+--------------------------------------------+---------------+
|C, R, S, D       |Parallel (nested)                           | ~ 32m         |
+-----------------+--------------------------------------------+---------------+
|C, R, S, D       |Parallel (nested), Precomp Integrals        | ~ 22m         |
+-----------------+--------------------------------------------+---------------+
|C, R, S, D       |Parallel (nested), Precomp Integrals, Sparse| ~ 4m          |
+-----------------+--------------------------------------------+---------------+


## Credit Surface

```{r echo = FALSE, warning=FALSE, message=FALSE, fig.align='center'}
library('plotly')

cs <- read.csv('DataWork/LeverageCycle/StochCSRealistic10.csv', header = FALSE)

fig1 <- plot_ly(cs, x = ~V1, y = ~V2, z = ~V3, marker = list(color = ~V3, colorscale = 'Portland', showscale = TRUE, size = 3),
               hovertemplate = "FICO: %{x}<extra></extra> <br>LTV: %{y}<br> m: %{z:.2f}%<br>")

fig1 <- fig1 %>% add_markers()
fig1 <- fig1 %>% layout(scene = list(xaxis = list(title = 'FICO'),
                                   yaxis = list(title = 'OLTV'),
                                   zaxis = list(title = 'mortgage rate (%)')))
fig1

```


## Demand Curve: In-Sync

```{r echo = FALSE, warning=FALSE, message=FALSE, fig.align='center'}

demandCurve <- read.csv(file = 'DataWork/LeverageCycle/DemandCurveSynced.csv')

f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "housing quantity (q)",
  titlefont = f
)
y <- list(
  title = "house price (p)",
  titlefont = f
)

fig2 <- plot_ly(data = demandCurve, x = ~q, y = ~p,
                marker = list(size = 10,
                             color = 'rgba(255, 182, 193, .9)',
                             line = list(color = 'rgba(152, 0, 0, .8)',
                                         width = 2)))
fig2 <- fig2 %>% add_trace(data = demandCurve, x = ~q, y = ~p, type="scatter", mode = "lines")

fig2 <- fig2 %>% layout(xaxis = x, yaxis = y)

fig2



```


## Next Steps

1. Sparse grid implementation; interpolation
2. University HPC cluster (Grace)
